---
title: "well-annotated/commented"
author: "Emiley Garcia-Zych"
date: "2023-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## challenge 1
## create a vector of 1 word character strings comprising the first line of the novel Ulysses. "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed." 



## extract every third element of this vector. 


## the following function, gsub("[[:punct:]]","", *string*), applied to *string* will remove punctuation and special characters from that string. 



## challenge 2
## given the matrix, m, above, extract the 2nd,3rd, and 6th columns and assign them to the variable x. 


## given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x.


## given the matrix, m, above extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x. 




## challenge 3
## construct a 4-dimensional, 400 element ray (5x5x4x4) named "a" consisting of the numbers 400 to 1(i.e., a descending series)




## Given this matrix, what would the following return? 
## a[1,1,1,2]




## a[2,3,2,]




## a[1:5,1:5,3,3]






## challenge 4
## create a list representing this (simplified) primate taxonomy: 











## challenge 5
## store the following numbers as a 5 x 3 matrix, called "m3": 3, 0, 1, 23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2







## then, do the following
## coerce the matrix to a data frame. 



## as a data frame, coerce the second column to be logical-valued



## as a data frame, coerce the third column to be factor-valued



## when you are done, use the str() command to show the data type for each variable in your dataframe. 



