---
title: "well-annotated/commented"
author: "Emiley Garcia-Zych"
date: "2023-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Challenge 1

a)  Create a vector of 1 word character strings comprising the first line of the novel Ulysses. "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."

```{r}
ulysses_quote
```

```{r}
ulysses_quote_vector<- c(strsplit(ulysses_quote, " ")[[1]])
ulysses_quote_vector
```

b)  Extract every third element of this vector.

```{r}
##count <-0
##for (i in ulysses_quote_vector){
##  count = count + 1
##  if (count == 3){
##    print(i)
##    count = 0
##  }
##}

ulysses_quote_vector [seq(from = 0, to = length(ulysses_quote_vector), by = 3)]
```

c)  the following function, gsub("[[:punct:]]","", *string*), applied to *string* will remove punctuation and special characters from that string.

```{r}
ulysses_quote1 <- gsub("[[:punct:]]", "", ulysses_quote)
ulysses_quote1
```

```{r}
ulysses_quote_vector1<- c(strsplit(ulysses_quote1, " ")[[1]])
ulysses_quote_vector1
```

```{r}
ulysses_quote_vector1 [seq(from = 0, to = length(ulysses_quote_vector1), by = 3)]
```

## Challenge 2

*matrix example*

```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```

*...extracting the elements in row 4, column 5, and assign to the new variable "x" as follows:*

```{r}
x <- m[4, 5]
x
```

*...extracting an entire row*

```{r}
y <- m[6,]
y
```

*...extracting an entire column*

```{r}
z <- m[,3]
z
```

a)  Given the matrix, m, above, extract the 2nd,3rd, and 6th columns and assign them to the variable x.

```{r}
x <- matrix(data = c(m[,2],m[,3],m[,6]),nrow = 8, ncol = 3, byrow = TRUE)
x
```

b)  Given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x.

```{r}
x <- matrix(data = c(m[6,],m[7,],m[8,]), nrow = 3, ncol = 10, byrow = TRUE)
x
```

c)  Given the matrix, m, above extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x.

```{r}
x <- matrix(data = c(m[2:6,2:9]), nrow = 5, ncol = 8, byrow = FALSE)
x
```

## Challenge 3

a)  Construct a 4-dimensional, 400 element ray (5x5x4x4) named "a" consisting of the numbers 400 to 1(i.e., a descending series)

```{r}
a <- array(data = 400:1, dim = c(5,5,4,4))
a
```

b)  Given this matrix, what would the following return?

a[1,1,1,2]

```{r}
a[1,1,1,2]
```

a[2,3,2,]

```{r}
a[2,3,2,]
```

a[1:5,1:5,3,3]

```{r}
a[1:5,1:5,3,3]
```

## Challenge 4

a)  Create a list representing this (simplified) primate taxonomy:

```{r}
## Haplorhini 
##Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
##Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- c("Platyrrhini", "Catarrhini")
Tarsioidea <- c("Tarsiidae")
Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- list("Anthropoidea", "Tarsioidea")
Haplorhini

Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae", "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Strepsirhini
```

```{r}
library(data.tree)
primates <- Node$new("Primates")
  haplorhini <- primates$AddChild ("Haplorhini")
    anthropoidea <- haplorhini$AddChild ("Anthropoidea")
      platyrrhini <- anthropoidea$AddChild ("Platyrrhini")
        cebidae <- platyrrhini$AddChild ("Cebidae")
        atelidae <- platyrrhini$AddChild ("Atelidae")
        pitheciidae <- platyrrhini$AddChild ("Pitheciidae")
      catarrhini <- anthropoidea$AddChild ("Catarrhini")
        cercopithecidae <- catarrhini$AddChild ("Cercopithecidae")
        hylobatidae <- catarrhini$AddChild ("Hylobatidae")
        hominidae <- catarrhini$AddChild ("Hominidae")
    tarsioidea <- haplorhini$AddChild ("Tarsioidea")
        tarsiidae <- tarsioidea$AddChild ("Tarsiidae")
  strepsirhini <- primates$AddChild ("Strepsirhini")
    lorisoidea <- strepsirhini$AddChild ("Lorisoidea")
      lorisdae <- lorisoidea$AddChild ("Lorisdae")
      galagidae <- lorisoidea$AddChild ("Galagidae")
    lemuroidea <- strepsirhini$AddChild ("Lemuroidea")
      cheirogaleidae <- lemuroidea$AddChild ("Cheirogaleidae")
      lepilemuridae <- lemuroidea$AddChild ("Lepilemuridae")
      indriidae <- lemuroidea$AddChild ("Indriidae")
      lemuridae <- lemuroidea$AddChild ("Lemuridae")
      daubentoniidae <- lemuroidea$AddChild ("Daubentoniidae")
print(primates)
```

## Challenge 5

a)  store the following numbers as a 5 x 3 matrix, called "m3": 3, 0, 1, 23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2

```{r}
m2 <- c(3, 0, 1, 23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2)
m3 <- matrix(data = m2, nrow = 5, ncol = 3, byrow = TRUE)
m3
```

b)  ...coerce the matrix to a data frame.

```{r}
df <- as.data.frame(m3)
str(df)
```

c)  ...as a data frame, coerce the second column to be logical-valued

```{r}
new_df_x2 <- as.logical(df$V2)
new_df_x2
```

d)  ...as a data frame, coerce the third column to be factor-valued

```{r}
new_df_x3 <- as.factor(df$V3)
new_df_x3
```

e)  When you are done, use the str() command to show the data type for each variable in your dataframe.

```{r}
str(df)
str(new_df_x2)
str(new_df_x3)
```
